---
title: "Cross validation"
author: "Aiying Huang"
date: "2023-11-09"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal()+theme(legend.position="bottom"))

options(
  ggplot2.continuous.color="viridis",
  ggplot2.continuous.fill="viridis"
)
scale_colour_discrete=scale_color_viridis_d
scale_fill_discrete=scale_fill_viridis_d

set.seed(1)
```

## Nonlinear data and CV

```{r}
nonlin_df=
  tibble(
    id=1:100,
    x=runif(100,0,1),
    y=1-10*(x -.3)^2+rnorm(100,0,0.3)
  )

nonlin_df|>
  ggplot(aes(x=x,y=y))+
  geom_point()
```

DO the train/test split

```{r}
train_df=sample_n(nonlin_df,80)
test_df=anti_join(nonlin_df,train_df,by="id")
```

```{r}
train_df|>
  ggplot(aes(x=x,y=y))+
  geom_point()+
  geom_point(data=test_df,color="red")
```

```{r}
linear_mod=lm(y~x,data=train_df)
```

quick visualization of the linear model

```{r}
train_df|>
  modelr::add_predictions(linear_mod)|>
  ggplot(aes(x=x,y=y))+
  geom_point()+
  geom_line(aes(y=pred))
```

load key packages.

```{r}
library(tidyverse)
library(modelr)

set.seed(1)
```

```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = mgcv::gam(y ~ s(x), data = train_df)
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

```{r}
train_df |> 
  modelr::add_predictions(smooth_mod) |> 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
train_df |> 
  modelr::add_predictions(wiggly_mod) |> 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred))

```

```{r}
train_df |> 
   modelr::gather_predictions(linear_mod, smooth_mod, wiggly_mod) |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

The modelr has other outcome measures – RMSE is the most common, but median absolute deviation is pretty common as well.

The RMSEs are suggestive that both nonlinear models work better than the linear model, and that the smooth fit is better than the wiggly fit. However, to get a sense of model stability we really need to iterate this whole process. Of course, this could be done using loops but that’s a hassle …

```{r}
cv_df = 
  crossv_mc(nonlin_df, 100) 

cv_df =
  cv_df |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```


```{r}
library(tidyverse)
library(modelr)
library(mgcv)
set.seed(1)
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )

nonlin_df |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point()

cv_df = 
  crossv_mc(nonlin_df, 100) 

```

```{r}
cv_df =
  cv_df |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```


```{r}
cv_df = 
  cv_df |> 
  mutate(
    linear_mod  = map(train, \(df) lm(y ~ x, data = df)),
    smooth_mod  = map(train, \(df) mgcv::gam(y ~ s(x), data =df)),
    wiggly_mod  = map(train, \(df) gam(y ~ s(x, k = 30), sp = 10e-6, data = df))) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),#clean the output
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)),
    rmse_wiggly = map2_dbl(wiggly_mod, test, ~rmse(model = .x, data = .y)))
```

```{r}
cv_df = 
  cv_df |> 
  mutate(
    linear_mod  = map(train, ~lm(y ~ x, data = .x)),
    smooth_mod  = map(train, ~mgcv::gam(y ~ s(x), data = .x)),
    wiggly_mod  = map(train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),#clean the output
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)),
    rmse_wiggly = map2_dbl(wiggly_mod, test, ~rmse(model = .x, data = .y)))
```

```{r}
cv_df
```

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```





